biclaR: Estimating the socio-environmental impacts of car substitution by bicycle and public transit using open tools

### 2023-09-27 paper/PaperTRA/PaperTRA.Rmd
paper table 1

header-includes:
  # - \usepackage[hidelinks]{hyperref} #this is to remove the annoying boxes. but how to make them blue?
  - \usepackage{hyperref}
  - \hypersetup{colorlinks = TRUE,  urlcolor = blue, linkcolor = blue, citecolor = blue}
   #Colours links instead of ugly boxes
  #Colour for external hyperlinks
  #Colour of internal links
  #Colour of citations

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      out.extra = "")  # This forces knitr to label all figures.
```

@@

A Tabela 4 apresenta o número de viagens diárias do cenário base e potenciais, passíveis de serem realizadas até 5 km em bicicleta em complemento com TP, por tipo de percurso.
Para as redes segura e direta, as viagens do cenário base até 5 km em bicicleta + TP correspondem a 20% do total de viagens reportadas no inquérito.

Este cenário valoriza a utilização da bicicleta como complemento ao transporte público, com potencial de aumentar as viagens em TP realizadas na área metropolitana de Lisboa em até 12% (acrescentando às 825 mil do IMob 2017).
Estes resultados sugerem que o potencial de transferência de viagens em automóvel para bicicleta + TP poderá ser quase ou tão importante quanto a transferência para viagens utilizando apenas a bicicleta.

Para este cenário foram estimados os impactes ambientais de transferência do automóvel para os transportes públicos, para além dos impactes ambientais e sociais da transferência para a bicicleta

Foram então calculadas as emissões dos poluentes para cada viagem substituída, tendo em conta o modo de transporte, distâncias e velocidade.

Para esse cenário com essas características, é estimada uma poupança de 8 702 toneladas de CO~2~eq por via da transferência de viagens motorizadas com combustíveis fósseis e eletricidade, numa perspetiva de ciclo de vida da produção de eletricidade.

\[Tabela 13\]

@@

```{r summary1, out.width="100%", fig.cap="Summary of the cycling potencial of intermodality scenario."}
knitr::include_graphics("img/table1.png", error = FALSE)
```

@@

@@

By making the research process publicly accessible in a code repository, this study enables the replication of similar estimates for socio-environmental impacts resulting from a modal shift from cars to bicycles + PT in other metropolitan areas.

@@

@@

The provided information on socio-economic benefits can support policymakers in prioritizing interventions to reduce the reliance on individual motorized transportation and effectively communicate their decisions.

>>>>

header-includes:
  # - \usepackage[hidelinks]{hyperref} #this is to remove the annoying boxes. but how to make them blue?
  - \usepackage{booktabs}
  - \usepackage{siunitx}
  - \usepackage{hyperref}
  - \hypersetup{colorlinks = TRUE,  urlcolor = blue, linkcolor = blue, citecolor = blue}
   #Colours links instead of ugly boxes
  #Colour for external hyperlinks
  #Colour of internal links
  #Colour of citations

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      out.extra = "")  # This forces knitr to label all figures.
library(dplyr)
```

@@

Table \ref{tab:summary1} presents the LMA total daily trips, the cycling combined with PT trips in the baseline scenario and corresponding potential trips to achieve the national strategy targets (4% and 10%), for different route profiles. 
For the cycling legs of the journey (first and last legs), the environmental avoided emmissions and monetized socio-environment benefits are also presented, resulting from replacing car trips with cycling.

For both *direct* and *safe* route profiles, 20% of the daily car trips have the potential to be replaced by a combination of PT and cycling (up to 5 km on bike).

@@

```{r summary1old, eval=FALSE, out.width="100%", fig.cap="Summary of the cycling potencial of intermodality scenario."}
knitr::include_graphics("img/table1.png", error = FALSE)
```
```{r summary1}
summary1 = readRDS("load/summary1.Rds")

summary1 = summary1 %>%
  mutate(LTS = recode(LTS, `3` = "safe", `4` = "direct")) |> 
  mutate(Estrategia = paste0(Estrategia*100,"%"))

knitr::kable(summary1,
      digits = 0,
      format.args = list(big.mark = " "),
      col.names = c("Target", "Routing", "Total trips",
                    "Baseline Cycling + PT", "Potencial Cycling + PT",
                    "Avoided CO2eq (ton/yr)", "SE Benefits for 10 years (thousand €)"),
      caption = "\\label{summary1old}Summary of the cycling potencial of intermodality scenario.",
      format="latex",
      booktabs = TRUE) |> 
  kableExtra::column_spec(column = c(4:7), width = "6em")
  # kableExtra::kable_styling(full_width = TRUE)

This scenario unveils the potential of bicycles as a complementary mode of PT, with the prospective to uptake the number of PT trips within the LMA area by as much as 12% (in addition to the 825,000 reported in the mobility survey).
These findings indicate that transferring car trips to a combination of bicycle and PT could be nearly as substantial, if not equally significant, as the shift towards bicycle-only trips.

<!-- this co2 needs to be added to the second table and then discuss -->
We estimate a reduction of 8,702 tons of CO2eq emissions by substituting motorized trips relying on fossil fuels and electricity. These estimations used an electricity production life cycle approach.

impactes ambientais de transferência do automóvel para os transportes públicos (tabela 5), para além dos impactes ambientais e sociais da transferência para a bicicleta (tabela 4)

@@

The estimated socio-economic impacts stemming from the avoided pollutants and greenhouse gases vary from €1.4 million to €3.5 million. Our findings suggest that a bicycle and PT combination could viably replace 20% of current trips, with an additional 12% of PT journeys prone to further substitution. The adoption of cycling for first-and-last-mile travel can curtail annual CO2 equivalent emissions by 6,000 to 15,000 tons per day. Over a decade, the resulting socio-environmental benefits could accrue to an estimated range of €230 million to €590 million, contingent upon the targeted cycling modalities. Regarding the PT segment, the shift from private car usage would lead to the mitigation of CO2 equivalent emissions to 8,500 to 20,800 tons annually, translating to a monetary value of €1.4 million to €3.5 million over a decade. Trains exhibit the highest potential for substitution, accounting for 88% of this benefit.

@@

By making the research process publicly accessible in a code repository, this study enables the replication of similar estimates for socio-environmental impacts resulting from a modal shift from car to bicycle in combination with PT in other metropolitan areas.

@@

Different routing profiles for different city cycling maturity levels [@felix2017].

@@

The provided information on socio-economic benefits can support policymakers in prioritizing interventions to reduce the reliance on individual motorized transportation and better communicate their decisions.

### 2023-09-27 paper/PaperTRA/PaperTRA.tex
paper table 1

\usepackage{hyperref}
\hypersetup{colorlinks = TRUE,  urlcolor = blue, linkcolor = blue, citecolor = blue}

@@

A Tabela 4 apresenta o número de viagens diárias do cenário base e
potenciais, passíveis de serem realizadas até 5 km em bicicleta em
complemento com TP, por tipo de percurso. Para as redes segura e direta,
as viagens do cenário base até 5 km em bicicleta + TP correspondem a
20\% do total de viagens reportadas no inquérito.

Este cenário valoriza a utilização da bicicleta como complemento ao
transporte público, com potencial de aumentar as viagens em TP
realizadas na área metropolitana de Lisboa em até 12\% (acrescentando às
825 mil do IMob 2017). Estes resultados sugerem que o potencial de
transferência de viagens em automóvel para bicicleta + TP poderá ser
quase ou tão importante quanto a transferência para viagens utilizando
apenas a bicicleta.

Para este cenário foram estimados os impactes ambientais de
transferência do automóvel para os transportes públicos, para além dos
impactes ambientais e sociais da transferência para a bicicleta

Foram então calculadas as emissões dos poluentes para cada viagem
substituída, tendo em conta o modo de transporte, distâncias e
velocidade.

Para esse cenário com essas características, é estimada uma poupança de
8 702 toneladas de CO\textsubscript{2}eq por via da transferência de
viagens motorizadas com combustíveis fósseis e eletricidade, numa
perspetiva de ciclo de vida da produção de eletricidade.

@@

\[Tabela 13\]

\begin{figure}
\includegraphics[width=1\linewidth,]{img/table1} \caption{Summary of the cycling potencial of intermodality scenario.}\label{fig:summary1}
\end{figure}

@@

@@

@@

>>>>

\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{hyperref}
\hypersetup{colorlinks = TRUE,  urlcolor = blue, linkcolor = blue, citecolor = blue}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

Table \ref{tab:summary1} presents the LMA total daily trips, the cycling
combined with PT trips in the baseline scenario and corresponding
potential trips to achieve the national strategy targets (4\% and 10\%),
for different route profiles. For the cycling legs of the journey (first
and last legs), the environmental avoided emmissions and monetized
socio-environment benefits are also presented, resulting from replacing
car trips with cycling.

For both \emph{direct} and \emph{safe} route profiles, 20\% of the daily
car trips have the potential to be replaced by a combination of PT and
cycling (up to 5 km on bike).

\begin{table}

\caption{\label{tab:summary1}\label{summary1old}Summary of the cycling potencial of intermodality scenario.}
\centering
\begin{tabular}[t]{llr>{\raggedleft\arraybackslash}p{6em}>{\raggedleft\arraybackslash}p{6em}>{\raggedleft\arraybackslash}p{6em}>{\raggedleft\arraybackslash}p{6em}}
\toprule
Target & Routing & Total trips & Baseline Cycling + PT & Potencial Cycling + PT & Avoided CO2eq (ton/yr) & SE Benefits for 10 years (thousand €)\\
\midrule
4\% & safe & 1 077 028 & 4 624 & 40 770 & 5 920 & 230 270\\
4\% & direct & 1 001 761 & 4 547 & 37 889 & 6 011 & 223 720\\
10\% & safe & 1 077 028 & 4 624 & 104 647 & 15 192 & 591 790\\
10\% & direct & 1 001 761 & 4 547 & 97 218 & 15 414 & 574 200\\
\bottomrule
\end{tabular}
\end{table}

This scenario unveils the potential of bicycles as a complementary mode
of PT, with the prospective to uptake the number of PT trips within the
LMA area by as much as 12\% (in addition to the 825,000 reported in the
mobility survey). These findings indicate that transferring car trips to
a combination of bicycle and PT could be nearly as substantial, if not
equally significant, as the shift towards bicycle-only trips.

@@

We estimate a reduction of 8,702 tons of CO2eq emissions by substituting
motorized trips relying on fossil fuels and electricity. These
estimations used an electricity production life cycle approach.

impactes ambientais de transferência do automóvel para os transportes
públicos (tabela 5), para além dos impactes ambientais e sociais da
transferência para a bicicleta (tabela 4)

@@

The estimated socio-economic impacts stemming from the avoided
pollutants and greenhouse gases vary from €1.4 million to €3.5 million.
Our findings suggest that a bicycle and PT combination could viably
replace 20\% of current trips, with an additional 12\% of PT journeys
prone to further substitution. The adoption of cycling for
first-and-last-mile travel can curtail annual CO2 equivalent emissions
by 6,000 to 15,000 tons per day. Over a decade, the resulting
socio-environmental benefits could accrue to an estimated range of €230
million to €590 million, contingent upon the targeted cycling
modalities. Regarding the PT segment, the shift from private car usage
would lead to the mitigation of CO2 equivalent emissions to 8,500 to
20,800 tons annually, translating to a monetary value of €1.4 million to
€3.5 million over a decade. Trains exhibit the highest potential for
substitution, accounting for 88\% of this benefit.

@@

Different routing profiles for different city cycling maturity levels
{[}19{]}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-felix2017}{}}%
\CSLLeftMargin{19. }%
\CSLRightInline{Félix, R., Moura, F., Clifton, K.J.: Typologies of urban
cyclists: Review of market segmentation methods for planning practice.
Transportation Research Record. 2662, 125--133 (2017).
https://doi.org/\href{https://doi.org/10.3141/2662-14}{10.3141/2662-14}.}
